\vspace{-2cm} Government websites as data: \\ A methodological pipeline with application to the websites of municipalities in the United States

### 2019-05-08 paper/manuscript_paNote.tex
first pass at pipeline section

%mportant to our methodological objectives, research analyzing and improving the scraping, pre-processing and text analysis pipeline that is applicable to government websites is still in its infancy. \cite{Eschenfelder2002} provide something of an overview of how federal websites should be assessed from an e-governance point of view, but they largely focus on the substantive criteria that should be fulfilled, rather than the technical aspects of website acquisition and analysis. %In what follows we first define the target dataset---the textual contents of websites of United States municipalities, along with associated metadata on the municipalities and their governments. We then define a pipeline for data collection and analysis that includes methods to access government website URLs, scrape their raw contents from the World Wide Web, gather plain text from the website files, and identify boilerplate text within the plain text contents. Lastly, we illustrate the analysis of municipal government website text by exploring the relationship between the city mayors' party affiliations and the topical contents of the websites.

@@

\section{The Web to Text Pipeline}

Once we have gathered the website files we need to pre-process the data \citep{denny2018text}. In this section, we describe our pre-processing pipeline, with which we take an archive of website files, and output a corpus of formatted plain text documents that are suitable for comparative analysis with text as data methods. In this pipeline, we address two methodological challenges. First, though they contain significant amounts of text, websites are not comprised of clean plain text files. Rather, the files available at websites are of multiple types, including HTML, PDF, word processor, plain text, and image files. The first step in the methodological pipeline is aimed simply at extracting clean plain text from this heterogeneous file base. The second step in our methodological pipeline is to process the text to remove language that is effective at differentiating one website from another but is uninformative regarding policy or political differences between governments. Some of the steps we take in this processing pipeline are universally applicable in the analysis of textual data, and some of them are most appropriate for the particular type of text analysis that we apply to this data---statistical topic modeling. We will clarify this distinction as we describe steps in our pipeline.

\subsection{Site to Text Conversion}
The format of a file has a major impact on whether and how textual data can be extracted from a document. For all text analysis projects, researchers need to consider file formats. For the most part, the file type of a document can be correctly determined through the filename ending---its extension. However, there are exceptions to this, which, if ignored, can lead to large amounts of improperly formatted text, arising from incorrectly converted documents, which leads to a general decrease in the amount of usable data. Two issues, in particular, need to be addressed: One, HTML files on city websites frequently do not have an ending but are still perfectly readable if correctly identified as such. Second, some documents contain the incorrect file ending. For example, we found thousands of documents that ended in .html, when they were actually PDFs. To accurately assess their type, we rely on the R package \texttt{wand} \citep{wand}, which is an R interface to the Unix library \texttt{libmagic} \citep{darwin2008libmagic}, which determines the type of a file on the basis of its file signature - or ``magic number''. This short sequence of bytes at the start (and sometimes end) of files is unique for each file type and therefore allows its correct identification through computer forensics tools such as \texttt{libmagic}. Files with incorrect endings are subsequently renamed.

\subsubsection{Extracting Text from HTML}
The HTML files that websites are comprised of contain a large amount of useful information, but also completely irrelevant text such as menus and navigational elements. We leverage methods developed in the information retrieval literature to deal with this problem. These so-called 'boilerplate detection' tools rely on both structural features, such as HTML tags, as well as text statistics such as word and sentence length to estimate whether a given portion of an HTML file is useful. We rely on the method described in \cite{Kohlschuetter2010}, which has been published in a reputable conference, is widely cited and can easily be implemented through the R package \texttt{boilerpipeR}.

\subsubsection{Extracting Text from PDF, XML, DOC, DOCX and TXT}
Other files are read in through the \texttt{readtext} R package\footnote{\texttt{readtext} determines a document's type solely through its ending -- so the conversion described above is strictly necessary for the package to work correctly.} \citep{readtext}, which is a wrapper for a set of parsers.\footnote{We have also experimented with several Unix-based alternatives, but found that they largely led to the same results as \texttt{readtext}.}\footnote{\texttt{readtext} also contains an HTML parser, but this tool simply extracts all text and is therefore inferior to boilerpipe.} The breakdown of the files by type is given in Table \ref{tab:filetypes}. The most frequent file types are HTML and PDF, from which we are able to extract a substantial amount of usable text. Files of type XML, DOC, TXT, and DOCX, also occur regularly in our corpus and offer a considerable volume of textual data.

@@

We then take several steps to pre-process the data as required for the subsequent analysis. Pre-processing choices should be contingent on the analysis being conducted with the text later and can have significant effects on the outcomes of an analysis \citep{denny2018text}. The type of text analysis we conduct---topic modeling---requires that the words in the document be meaningful and interpretable, and does not make use of the sequence of words within a document (i.e., is a bag-of-words method). %With the exception of file type conversion and boilerplate removal, the pre-processing steps taken here should therefore not be seen as universally applicable to all analyses of government website data.

For preprocessing, we rely on a set of tools which are used much more commonly in natural langauge processing than political methodology -- part-of-speech (POS) tagging and named entity recognition (NER), as these procedures perform very well in separating substantively interesting information from any remaining boilerplate. Names of city officials and citizen petitioners feature frequently in city documents. The same is true for streets, locations and not least of all, the city itself. Since individual names recur at a much higher rate within a city than across the entire corpus, this causes a topic model to cluster its topics by city. As these names convey no substantive information, we rely on NER to remove them. Specifically we use \texttt{spacyR} to detect and remove all entities except X, Y and Z, as well. Similarly we remove words on the basis of their POS-tags, retaining only nouns (by far the most informative category), verbs, adjectives. Furthermore, we keep proper nouns that also occur as nouns -- this removes names, but retains titles such as ``Police Chief'' which can appear as proper nouns if they are followed by a name.

We also conduct lemmatization to reduce words to their basic form. Lemmatization is similar to stemming but works in a somewhat more sophisticated manner by taking grammar and surrounding words into account to identify the dictionary form of a word. For example, the lemma of the word ``lemmatization'' would be ``lemmatize'', whereas most stemmers would simply chop off the ending, which would yield ``lemmatiz''. Thus, lemmatization makes the results more easily comprehensible.

Furthermore, we remove words with less than three characters (these are usually artifacts from improperly encoded documents and faulty or impartial optical character recognition), stopwords and non-English words (using the R package \texttt{Hunspell})\footnote{As we are using \texttt{spacyR} with an English tagger, non-English words are erroneously categorized as (proper) nouns.}.

A final and crucial step is the removal of duplicate documents, which occur very frequently on websites. In addition to their primary purpose, the previous preprocessing steps also help in stripping otherwise identical documents of information that makes them unique -- such as names and dates -- thus facilitating their deletion. The removal of duplicate information also solves a problem posed by a recurring feature of municipal websites -- city calendars. For most days in most cities, these calendars tend to be empty, but nevertheless contain a template, which would lead to the creation of a document in our corpus. These short and repetitive texts would pose severe problems to a topic model. Our approach preserves the valuable information contained within these calendars when cities do make use of them, but removes them when they offer no substantive data.

The text documents are converted to UTF-8 character encoding and then stripped of dates, punctuation, numbers, and words connected by underscores. At this point, the documents of one city still closely resemble one another in the form of boilerplate content, be it website elements (i.e. "You are here", "Home", "Directory" etc.) in HTML documents, or commonly used forms or phrases in pdf, doc and docx files. This is an issue, because this boilerplate content causes the results of analyzing this data with text analysis methods to characterize documents primarily by the cities from which they originate (through their unique boilerplate structure, e.g. a menu with certain terms repeated on every site of the domain), and not the substantive features of their contents. Boilerplate removal is a useful step in many forms of text analysis, as the analysis is focused in on text that varies above and beyond a standard template for textual content. Our solution to this problem is described in more detail in Section \ref{boilerplate}.

The last round of preprocessing is intended to remove everything from the file that is not an English word. This step is tailored to our intention to use the text for topic modeling. The final preprocessing round includes setting every character to lowercase, as well as the removal of bullet points which frequently occur in HTML documents, extraneous whitespace, XML documents mislabeled as HTML files, and empty documents. Furthermore, some documents contain gibberish, often as a result of faulty or impartial optical character recognition applied to text that was produced through a non-machine-readable medium. To combat this problem, we employ two solutions. One, we use spellchecking, implemented through the \texttt{hunspell} R package \citep{hunspell}, to remove all non-English words.\footnote{Some of the cities, for example, Los Angeles, do contain a sizable proportion of Spanish content. The analysis of this content is beyond the scope of this paper but could be explored in future work, for example using methods of text processing that are applicable to multilingual corpora \citep{lucas2015computer}. } However, \texttt{hunspell} does not cover everything, either because some tokens are not actual words (for example artifacts from defective encoding), or because random sequences of characters just so happen to form words that exist in a dictionary (for example "eh" or "duh"). Since we rely on a bag-of-words model in which syntax does not matter, we can ameliorate these problems by removing all text except for whitespaces and the characters that appear in the English alphabet. Since a lot of the nonsensical text tends to be quite repetitive, we also delete all documents in which the proportion of unique to the total number of tokens is less than 0.15. Furthermore, \texttt{hunspell} does not spellcheck individual characters or two-character words, so we remove these token types entirely. Since these pre-processing steps reduce documents which are largely unsuitable to only a few tokens (i.e., word occurrences), we also remove all remaining documents containing less than 50 tokens. Finally, to remove words that are extremely rare (which also has the advantage of eliminating any remaining oddities) and thus add nothing substantive\footnote{Topic models essentially do not pick up on extremely rare words, so their inclusion is a waste of computational resources. Removing them in this manner is also the default preprocessing choice in the \texttt{stm} package.} to our models while increasing their computational cost, we also discard any token types that occur in only one document.

\subsection{Boilerplate Removal}\label{boilerplate}
As noted above, city websites contain a large amount of text that is uninformative for its actual content, and therefore a hindrance to understanding through algorithmic text analysis. This is a common issue with textual data in which informative content is embedded in technically structured documents. See, e.g., \citet{burgess2016legislative,wilkerson2015tracing} and \citet{linder2018text} for examples of boilerplate removal in the analysis of legislative text.

>>>>

%Important to our methodological objectives, research analyzing and improving the scraping, pre-processing and text analysis pipeline that is applicable to government websites is still in its infancy. \cite{Eschenfelder2002} provide something of an overview of how federal websites should be assessed from an e-governance point of view, but they largely focus on the substantive criteria that should be fulfilled, rather than the technical aspects of website acquisition and analysis. %In what follows we first define the target dataset---the textual contents of websites of United States municipalities, along with associated metadata on the municipalities and their governments. We then define a pipeline for data collection and analysis that includes methods to access government website URLs, scrape their raw contents from the World Wide Web, gather plain text from the website files, and identify boilerplate text within the plain text contents. Lastly, we illustrate the analysis of municipal government website text by exploring the relationship between the city mayors' party affiliations and the topical contents of the websites.

@@

\section{The Web to Text Pipeline}
In this section, we describe our methodological pipeline, with which we take an archive of website files, and output a corpus of formatted plain text documents that are suitable for comparative analysis with text as data methods. Here, we address three methodological challenges.  First, though they contain significant amounts of text, websites are not comprised of clean plain text files. Rather, the files available at websites are of multiple types, including HTML, PDF, word processor, plain text, and image files. The first step is aimed at extracting clean plain text from this heterogeneous file base. The second step in our pipeline is to process the text to remove language that is effective at differentiating one website from another but is uninformative regarding policy or political differences between governments. Finally, these tools need to work consistently across all of the websites in our corpus, in spite of the fact that relevant information is stored and structured in different ways.%Some of the steps we take in this processing pipeline are universally applicable in the analysis of textual data, and some of them are most appropriate for the particular type of text analysis that we apply to this data---statistical topic modeling. We will clarify this distinction as we describe steps in our pipeline.

\subsection{Site to Text Conversion}
\subsubsection{File Type Detection}
The format of a file has a major impact on whether and how textual data can be extracted from a document. For the most part, the file type of a document can be correctly determined through the filename ending---its extension. However, there are exceptions to this, which, if ignored, can lead to large amounts of improperly formatted text, arising from incorrectly converted documents, which leads to a general decrease in the amount of usable data. Two issues, in particular, need to be addressed: One, HTML files on city websites frequently do not have an ending but are still perfectly readable if correctly identified as such. Second, some documents contain the incorrect file ending. For example, we found thousands of documents that ended in .html, when they were actually PDFs. To accurately assess their type, we rely on the R package \texttt{wand} \citep{wand}, which is an R interface to the Unix library \texttt{libmagic} \citep{darwin2008libmagic}, which determines the type of a file on the basis of its file signature - or ``magic number''. This short sequence of bytes at the start (and sometimes end) of files is unique for each file type and therefore allows its correct identification through computer forensics tools such as \texttt{libmagic}. Files with incorrect endings are subsequently renamed.

\subsubsection{Extracting Text from HTML}
The HTML files that websites are comprised of contain a large amount of useful information, but also completely irrelevant text such as menus, navigational elements and other boilerplate. In theory, HTML tags intended to mark text content  exist to prevent this problem. In practice, not every web developer adheres to this standard, and even when they do, there is a large degree of variation in the way it is implemented. Parsing any one website is fairly straightforward, but parsing the 283 different websites in our sample in a consistent and comparable manner makes a rule-based approach impossible.

We leverage methods developed in the information retrieval literature to deal with this problem. These boilerplate detection tools are classifiers which rely on both structural features, such as HTML tags, as well as text statistics such as word and sentence length to estimate whether a given portion of an HTML file is useful. We rely on the method described in \cite{Kohlschutter2010}, which has been published in a reputable conference, is widely cited and can easily be implemented through the R package \texttt{boilerpipeR}.

\subsubsection{Extracting Text from PDF, XML, DOC, DOCX and TXT}
Other files are read in through the \texttt{readtext} R package \citep{readtext}, which is a wrapper for a set of parsers.\footnote{\texttt{readtext} determines a document's type solely through its ending -- so the conversion described above is strictly necessary for the package to work correctly.}$^{,}$\footnote{We have also experimented with several Unix-based alternatives, but found that they largely led to the same results as \texttt{readtext}.}$^{,}$\footnote{\texttt{readtext} also contains an HTML parser, but this tool simply extracts all text and is therefore inferior to boilerpipe.} The breakdown of all files by type is given in Table \ref{tab:filetypes}. The most frequent file type besides HTML is PDF, from which we are able to extract a substantial amount of usable text. Files of type XML, DOC, TXT, and DOCX, also occur regularly in our corpus and offer a considerable volume of textual data.

@@

\subsection{Preprocessing}
Preprocessing is an important part of text-as-data research and choices made therein can have significant effects on the outcomes of an analysis \citep{denny2018text}. The challenge in conducting preprocessing for a comparative analysis of websites lies in the considerable variance between websites. Some of it is substantively informative and some of it is completely irrelevant. As an example of the latter, names of city officials and citizen petitioners feature frequently in city documents. The same is true for streets, locations and not least of all, the city itself. Since individual names recur at a much higher rate within a city than across the entire corpus, this would cause a topic model to cluster its topics by city. Consequently we require a tool which detects the signal in the noise and does so consistently for a discordant set of sources.

%The type of text analysis we conduct---topic modeling---requires that the words in the document be meaningful and interpretable, and does not make use of the sequence of words within a document (i.e., is a bag-of-words method). %With the exception of file type conversion and boilerplate removal, the pre-processing steps taken here should therefore not be seen as universally applicable to all analyses of government website data.

To this end, we turn to to an approach that is much more common in natural langauge processing than political methodology -- part-of-speech (POS) tagging and named entity recognition (NER). While not their original purpose, these procedures perform very well in separating the wheat from the chaff. As names convey no substantive information, we NER is used to detect to remove them.\footnote{We retain laws, nationalities or religious or political groups (which are politically salient with respect to immigration and identity) as well as works of art (which frequently feature statutes, plans, etc.)} Furthermore, we select words on the basis of their POS-tags, retaining only nouns (by far the most informative category), verbs, adjectives. Furthermore, we keep proper nouns that also occur as nouns -- this removes names, but retains titles such as ``Police Chief'' which can appear as proper nouns if they are followed by a name. Finally, we also conduct lemmatization to reduce words to their basic form.\footnote{Lemmatization is similar to stemming, but works in a somewhat more sophisticated manner by taking grammar and surrounding words into account to identify the dictionary form of a word. For example, the lemma of the word ``lemmatization'' would be ``lemmatize'', whereas most stemmers would simply chop off the ending, which would yield ``lemmatiz''. Thus, lemmatization makes the results more easily comprehensible.} POS-tagging, NER and lemmatization are all implemented through \texttt{spacyR}. This parsing-based approach is very effective in distilling a comparable corpus from a varied set of sources, as the grammatical rules of the English language remain their common denominator. To deal with any leftover issues, we remove words with less than three characters (these are usually artifacts from improperly encoded documents and faulty or impartial optical character recognition), stopwords and non-English words (using the R package \texttt{Hunspell})\footnote{As we are using \texttt{spacyR} with an English parser, non-English words tend to be erroneously categorized as (proper) nouns.}.

A final and crucial step is the removal of duplicate documents, which occur very frequently on websites. In addition to their primary purpose, the previous preprocessing steps also help in stripping otherwise identical documents of information that makes them unique -- such as names and dates -- thus facilitating their deletion. The removal of duplicate information also solves a problem posed by a recurring feature of municipal websites -- city calendars. For most days in most cities, these calendars tend to be empty, but nevertheless contain a template, which would lead to the creation of a document in our corpus. These short and repetitive -- but numerous -- texts would pose severe problems to a topic model. Our approach preserves the valuable information contained within these calendars when cities do make use of them, but removes them when they offer no substantive data.

%The text documents are converted to UTF-8 character encoding and then stripped of dates, punctuation, numbers, and words connected by underscores. At this point, the documents of one city still closely resemble one another in the form of boilerplate content, be it website elements (i.e. "You are here", "Home", "Directory" etc.) in HTML documents, or commonly used forms or phrases in pdf, doc and docx files. This is an issue, because this boilerplate content causes the results of analyzing this data with text analysis methods to characterize documents primarily by the cities from which they originate (through their unique boilerplate structure, e.g. a menu with certain terms repeated on every site of the domain), and not the substantive features of their contents. Boilerplate removal is a useful step in many forms of text analysis, as the analysis is focused in on text that varies above and beyond a standard template for textual content. Our solution to this problem is described in more detail in Section \ref{boilerplate}.

%The last round of preprocessing is intended to remove everything from the file that is not an English word. This step is tailored to our intention to use the text for topic modeling. The final preprocessing round includes setting every character to lowercase, as well as the removal of bullet points which frequently occur in HTML documents, extraneous whitespace, XML documents mislabeled as HTML files, and empty documents. Furthermore, some documents contain gibberish, often as a result of faulty or impartial optical character recognition applied to text that was produced through a non-machine-readable medium. To combat this problem, we employ two solutions. One, we use spellchecking, implemented through the \texttt{hunspell} R package \citep{hunspell}, to remove all non-English words.\footnote{Some of the cities, for example, Los Angeles, do contain a sizable proportion of Spanish content. The analysis of this content is beyond the scope of this paper but could be explored in future work, for example using methods of text processing that are applicable to multilingual corpora \citep{lucas2015computer}. } However, \texttt{hunspell} does not cover everything, either because some tokens are not actual words (for example artifacts from defective encoding), or because random sequences of characters just so happen to form words that exist in a dictionary (for example "eh" or "duh"). Since we rely on a bag-of-words model in which syntax does not matter, we can ameliorate these problems by removing all text except for whitespaces and the characters that appear in the English alphabet. Since a lot of the nonsensical text tends to be quite repetitive, we also delete all documents in which the proportion of unique to the total number of tokens is less than 0.15. Furthermore, \texttt{hunspell} does not spellcheck individual characters or two-character words, so we remove these token types entirely. Since these pre-processing steps reduce documents which are largely unsuitable to only a few tokens (i.e., word occurrences), we also remove all remaining documents containing less than 50 tokens. Finally, to remove words that are extremely rare (which also has the advantage of eliminating any remaining oddities) and thus add nothing substantive\footnote{Topic models essentially do not pick up on extremely rare words, so their inclusion is a waste of computational resources. Removing them in this manner is also the default preprocessing choice in the \texttt{stm} package.} to our models while increasing their computational cost, we also discard any token types that occur in only one document.

%\subsection{Boilerplate Removal}\label{boilerplate}
%As noted above, city websites contain a large amount of text that is uninformative for its actual content, and therefore a hindrance to understanding through algorithmic text analysis. This is a common issue with textual data in which informative content is embedded in technically structured documents. See, e.g., \citet{burgess2016legislative,wilkerson2015tracing} and \citet{linder2018text} for examples of boilerplate removal in the analysis of legislative text.
